---
title: Plot Confusion Matrix in R
author: Enrique Pérez Herrero
date: '2022-03-09'
slug: []
categories:
  - R
  - code
  - machine learning
tags:
  - matrix
  - plot
  - ML
featured:
  url: conf_matrix.png
  alt: ggplot2 confusion matrix
  caption: Naïve Bayes prediction confusion matrix
  credit: Data Energy
  previewOnly: false
---

Plotting the Confusion Matrix with `ggplot` and R

<!--more-->

The R package `caret` includes the `confusionMatrix` function that returns a very complete output.

```{r echo=TRUE, message=FALSE}
library(e1071)
library(caTools)
library(ggplot2)
library(scales)
library(caret)
```

We run the Naïve Bayes classification in the classical iris data set.

```{r}
iris$spl = caTools::sample.split(iris, SplitRatio = 0.8)
train <- subset(iris, iris$spl == TRUE)
test <- subset(iris, iris$spl == FALSE)

iris.nb <- naiveBayes(Species ~ ., data = train)

nb_train_predict <-
  predict(iris.nb, test[, names(test) != "Species"])

cfm <- confusionMatrix(nb_train_predict, test$Species)
cfm
```

To plot the obtained confusion matrix as a `ggplot` graphic, we use the following function:

```{r}
ggplotConfusionMatrix <- function(m){
  mytitle <- paste("Accuracy", percent_format()(m$overall[1]),
                   "Kappa", percent_format()(m$overall[2]))
  p <-
    ggplot(data = as.data.frame(m$table) ,
           aes(x = Reference, y = Prediction)) +
    geom_tile(aes(fill = log(Freq)), colour = "white") +
    scale_fill_gradient(low = "white", high = "steelblue") +
    geom_text(aes(x = Reference, y = Prediction, label = Freq)) +
    theme(legend.position = "none") +
    ggtitle(mytitle)
  return(p)
}
```

```{r include=FALSE}
ggplotConfusionMatrix(cfm)
```


# Links:

* StackOverflow: [How to produce a confusion matrix and find the misclassification rate of the Naïve Bayes Classifier?](https://stackoverflow.com/questions/46063234/how-to-produce-a-confusion-matrix-and-find-the-misclassification-rate-of-the-na%c3%af/46063613#46063613)
